---
title: "Lab 03 - Nobel laureates"
author: "Fanyi Zeng"
date: "01/30/22"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
```

```{r load-data, message=FALSE}
nobel <- read_csv("data/nobel.csv")
```

## Exercises

### Exercise 1

935 observations (rows) of 26 variables (columns)

```{r view}
view(nobel)
```

### Exercise 2

filter out recipients who have no country data, who are organizations, and who have passed away.

```{r filter}
nobel_living <- nobel %>%
  filter(!is.na(country),
         gender != "org",
         is.na(died_date))
view(nobel_living)
```

### Exercise 3

sort country where the recipients live at the time of receiving the prize to US vs. other

```{r us-based}
nobel_living <- nobel_living %>%
  mutate(country_us = if_else(country == "USA", "USA", "Other"))
```

filter out non-science prizes

```{r science}
nobel_living_science <- nobel_living %>%
  filter(category %in% c("Physics", "Medicine", "Chemistry", "Economics"))
```

create a faceted bar plot visualizing the relationship between the category of prize and whether the laureate was in the US when they won the nobel prize

```{r prize-and-us-based}
ggplot(data = nobel_living_science, aes(y = country_us)) +
  geom_bar() +
  facet_wrap(~category)

```

True, even by each prize category, most living laureates are from the US.

### Exercise 4

sort ppl into us-born vs. other
105 rows of us-based laureates

```{r us-born}
nobel_living_science <- nobel_living_science %>%
  mutate(born_us = if_else(born_country == "USA", "USA", "Other") )
nobel_living_science %>%
  filter(born_us == "USA")
```

### Exercise 5

add a second variable to your visualization from Exercise 3 based on whether the laureate was born in the US or not

```{r prize-and-us-born}
ggplot(data = nobel_living_science, aes(y = country_us, fill = born_us)) +
  geom_bar() +
  facet_wrap(~category) +
  labs(
    x = "Count",
    y = "Country of Residence",
    fill = "Country of Origin",
    title = "Residing Country of Living Laureates",
    subtitle = "By Country of Origin"
  )
```

Not true. Most USA-based laureates are born in  the USA.

### Exercise 6

The most common country of origin for US-based living laureates who are not born in the U.S. is the U.K.

```{r birth-country}
nobel_living_science %>%
  filter(country_us == "USA", born_us == "Other") %>%
  count(born_country) %>%
  arrange(desc(born_country))